---
title: "Untidy data"
output: github_document
author: "Stefanos Karageorgiou, Tianwei Liu, Manuel Gonzalez, Yilun Dong"
editor_options: 
  chunk_output_type: inline
---

## Setup - Load the data

```{r setup, include = FALSE}
library(dplyr)
library(ggplot2)
library(tidyr)
```


## Lego Sales Data

```{r message = FALSE}
sales <- readRDS("/Users/stefanoskarageorgiou/Desktop/lego_sales.rds")
```

### Part 1 - Tidying the data
Write up:

Part 1:<br />
We created a tidy data frame with the data cointained in the file `lego_sales.rds`, which is a nested list. We used commands from the package `dplyr` in order to map the variables in the list into columns. The result was a tibble of 620 observations, one for each purchase of a different Lego set made by a different costumer, and 16 columns. The first 6 columns correspond to variables from the costumer, and the last 10 are variables from the purchase. We decided to leave the variable `hobbies` as a list as we thought unnesting it would make the tibble extremely large and uneasy to explore. 


```{r}
TS = 
  tibble(sales) %>%
  hoist(
    sales,
    gender = "gender",
    first_name = "first_name",
    last_name = "last_name",
    age = "age",
    phone_number = "phone_number",
    hobbies = "hobbies", 
    purchases = "purchases"
  ) %>%
  unnest_longer("purchases") %>%
  hoist(
    purchases,
    SetID = "SetID",
    Number = "Number",
    Theme = "Theme",
    Subtheme ="Subtheme",
    Year = "Year",
    Name = "Name",
    Pieces = "Pieces",
    USPrice = "USPrice", 
    ImageURL = "ImageURL", 
    Quantity = "Quantity"
    )
```


```{r}
print(TS)
```

### Possible questions about data and appropriate answers:


1. What are the three most common first names of purchasers?<br />
Explanations:<br />
For this question, first we selected the variables `first_name` and `last_name`. We then selected the distinct rows, as they appeared one time for each purchase. Afterwards, we counted the appearances of every first name and then selected the top 5. As there where ties in the first and third possition, the result was a list of 7 names.  

Comment 1:<br />
Jacob and Michael are the most common names, as they appear four times each in the database. After those, there are five names with three appearances.


```{r}
TS %>%
  select(first_name,last_name)%>%
  distinct()%>%
  count(first_name) %>% 
  arrange(-n) %>%
  top_n(3,n) %>%
  print(n = Inf)
```
Comment 2:<br />
Here top_n(3,n) will seek for the top 3 values of the variable `n`, which corresponds to the number of times that a name appears. In our case the maximum value of `n` is 4, but it occurres in 2 names. As a result, top_n() will pledge 2 spots for those names, leaving room for one more name. It is only logical that it will look for names with the second highest frequency which are those with n=3, so it will fill the remaining spot with the name correspoding to this value of `n`. However there are 5 names with the same value, so top_n(3,n) will accept them all and will create extra spots for all of them. Thus, top_n(3,n) will result in 7 names. In the case where we want to observe all the names with the 3 highest frequencies we would need a different code, which is included as a comment below.

```{r}
# TS %>%
#  select(first_name,last_name)%>%
#  distinct()%>%
#  count(first_name) %>%
#  arrange(-n) %>%
#  filter(dense_rank(-n)<=3)%>%
#  print(n = Inf)
```


2. Which lego theme has made the most money for lego?<br />
Explanations:<br />
In this problem, we used the variables `USPrice`, `Quantity`, and `Theme`. Then we created a new variable named `Revenue` by multiplying `USPrice` and `Quantity`. Afterwards, we group by `Theme`, and summed `Total_Revenue` for each group.

Comment:<br />
Here we can see that the theme with the highest revenue is Star Wars with the total revenue of 4447.83 US Dollars for lego.

```{r}
TS %>% 
  select(USPrice,Quantity,Theme) %>% 
  arrange(Theme) %>% 
  mutate(Revenue=Quantity*USPrice) %>%
  group_by(Theme) %>% 
  summarise(Total_Revenue=sum(Revenue)) %>%
  arrange(-Total_Revenue) %>%
  top_n(1,Total_Revenue) %>%
  print(n = Inf)
```
3. Do men or women buy more lego sets (per person) on average?<br />
Explanations:<br />
Firstly, we selected the variables `first_name`, `last_name`, `gender` and `Quantity`. We then grouped by `first_name`, `last_name` and `gender` and summed over `Quantity` in order to obtain the total number of sets purchased by every costumer. Then we grouped by gender and obtained the average quantity for each group.    

Comment:<br />
Altough the difference is very small, women buy on average more Legos than man. Women buy 3.57 Legos per person on average, as men buy 3.56.

```{r}
TS %>%
  select(first_name,last_name,Quantity,gender) %>%
  group_by(first_name,last_name,gender) %>%
  summarise(TotalQuantity=sum(Quantity)) %>%
  group_by(gender) %>%
  summarise(avgquantity=mean(TotalQuantity))
```


4. What are the five most popular hobbies of lego purchasers?<br />
Explanations:<br />
To tackle this problem, we selected `first_name`, `last_name`, and `hobbies`. In each row, `hobbies` remained as a list, so the next step was to unnest it. Afterwards, we selected the distinct values in order to obtain the different clients with their corresponding hobbies. Finally, we counted over the hobbies, and selected the top 5. Once again there were ties, so the result includes 12 hobbies.   

Comment:<br />
We once again come across with a similar case of question 1. Here top_n() will give us the 2 hobbies with the highest frequency which is 6, then 2 with the second highest frequency. This leaves one empty spot for the hobbie with the third highest frequency. As there are 8 observations with the same frequency, R included all of them. We used the same approach as before. Hence, we included a relevant code below in case we want to observe all the hobbies with the 5 highest frequencies, that total 107.

```{r}
  TS %>%
   select(first_name,last_name,hobbies) %>%
   unnest_longer(hobbies) %>%
   distinct() %>%
   filter(!is.na(hobbies)) %>%
   count(hobbies) %>%
   arrange(-n) %>%
   top_n(5,n) %>%
   print(n = Inf)
```

The table above shows that animal fancy and Skateboarding are the most common hobbies, each one appearing 6 times. The second most popular ones are Kitesurfing and Sculling or Rowing, with 4. Finally, there are 8 hobbies with four appearances.  

If we wanted to obtain all the hobbies with the 5 highest number of appearances, our solution will be as follows.  

```{r}
# TS %>%
#  select(first_name,last_name,hobbies) %>%
#  unnest_longer(hobbies) %>%
#  distinct() %>%
#  filter(!is.na(hobbies)) %>%
#  count(hobbies) %>%
#  arrange(-n) %>%
#  filter(dense_rank(-n)<=5) %>%
#  print(n = Inf)
```

5. Which area code has spent the most money on legos?<br />
Explanations:<br />
After selecting the relevant variables for this problem (USPrice, Quantity and phone_number), the second task was to obtain the area code (the first three digits of the phone number), so we separated the phone number into three different parts and selected the first one. We also created a new variable named `Revenue` by multiplicating `Quantity` and `USPrice`. Afterwards, we grouped by area code and summed the revenues.  

Comment:<br />
The area code 956, which corresponds to Laredo, TX, was the one that spended the most on Legos.

```{r}
TS %>%
  select(USPrice,Quantity,phone_number) %>%
  separate(phone_number,
           into=c("Area_Code","Phone_number_Part_2","Phone_number_Part_3"),
           sep="-") %>%
  select(Area_Code,USPrice,Quantity) %>%
  filter(!is.na(Area_Code)) %>%
  mutate(Revenue=USPrice*Quantity) %>%
  group_by(Area_Code) %>%
  summarise(total=sum(Revenue)) %>%
  arrange(-total) %>%
  top_n(1,total) %>%
  print(n = Inf)
```

## GitHub dataset

```{r message = FALSE}
commits = readRDS("dplyr_2019_commits.rds")
```

### Part 1 - Tidying the data

Write-up for part 1:<br />
In this sector, we created a data frame called `df_commits` containing 6 columns, then printed out the first 10 rows of it. The columns in the data frame are listed as follows:<br />
`commit_sha` is the unique identifier of each commit.<br />
`author_name` is the login name (unique identifier) of each author.<br />
`file_name` is the name with the full path of each file as the unique identifier of files.<br />
`hour` is the time of each commit in hours.<br />
`message` is the comment of each commit.<br />
`change_in_lines` is the total number of changes, namely addtions plus deletions in the commit.<br />

To explain the creation process in detail, firstly we created the initial data frame. Next, we unnested the column named `files` and put different files from the same commit into different rows. We also created a column called `file_name` when unnesting. Afterwards, we modified `time` column to keep only the hour of commit and deleted irrelevant time elements. Finally, we selected the 6 useful columns from this data frame, printed the first 10 rows of it and proceeded to the second part of the problem.


```{r}
df_commits = 
          tibble(commits) %>%
          hoist(
            commits,
            commit_sha = "sha",
            author_name = c("author","login"),
            files = "files",
            time = c("commit", "author", "date"),
            message = c("commit", "message"),
            change_in_lines = c("stats", "total")
          ) %>%
          unnest_longer("files") %>%
          hoist(
            files,
            file_name = "filename"
          ) 
df_commits =   
  df_commits %>% separate(time,into=c("date","time"),sep='T') %>%
                  separate(time,into=c("hour","minute","second"),sep=":") %>%
  select(commit_sha, author_name, file_name, hour, message, change_in_lines)
```


```{r}
print(df_commits)
```


### Part 2 - Questions

1. Who are the top five contributors (in terms of the most commits) to dplyr in the last year?

Explanations to problem 1:<br />
In this problem, we firstly selected `commit_sha` and `author_name` from `df_commits`. Then we used `distinct()` to eliminate repeted `commit_sha` from the new data frame. This is necessary because one commit could consist of multiple files with the same `commit_sha` but different `file_name` were put into different rows. Next, we counted the number of commits made by each author and arranged the results in descending order. Finally, we applied `top_n()` to include tied values.<br />

```{r}
df_commits %>%
  select(commit_sha, author_name) %>%
  distinct() %>%
  group_by(author_name) %>%
  summarise(commits = n()) %>%
  arrange(-commits) %>%
  top_n(5, commits) %>%
  print(n = Inf)
```

2. Which four files have been modified in the most number of commits?

Explanations to problem 2:<br />
This problem is similar to the last one. Firstly, we selected `file_name` from `df_commits`. Next, we counted the appearances of each file and sorted the results in descending order. Again, we used `top_n()` to include tied values at the end.<br />

```{r}
df_commits %>%
  select(file_name) %>%
  group_by(file_name) %>%
  summarise(times_of_modified=n()) %>%
  arrange(-times_of_modified) %>%
  top_n(4, times_of_modified) %>%
  print(n = Inf)
```

3. When is the most active time of the day for development on dplyr? More specifically, what is the average number of total changes made per commit for each hour of the day?

Explanations to problem 3:<br />
In this problem, we selected `commit_sha`, `hour` and `change_in_lines` from `df_commits`. Then we removed comments including 'revdep' and 'Merge', grouped the data by hour, calculated the avegare numbers and put them in descending oreder. The answer can be read directly from the output.<br />

```{r}
df_commits %>%
  select(commit_sha,hour,change_in_lines, message) %>%
  distinct() %>%
  filter(grepl("revdep", message) == FALSE & grepl("Merge", message) == FALSE) %>%
  group_by(hour) %>%
  summarize(average_change_in_lines=mean(change_in_lines)) %>%
  arrange(-average_change_in_lines) %>%
  print(n = Inf)
```

4. Based on these data is there any "evidence" that commit message length is related to the complexity of the commit (as measured by the total number of changes in the commit)? Justify your answer.

Explanations to problem 4:<br />
To tackle this problem, we firstly selected `commit_sha`, `message` and `change_in_lines` from `df_commits` and created `df_4`. Then we used `distinct()` to delete repeated `commit_sha`. The reason for doing this has been stated in the write up for Problem 1. Afterwards, we applied two functions to examine the corelation between length of `message` and number of `change_in_lines`. The result from Pearson test illustrated that there was very little evidence of linear relationship between these two variables. However, Spearman test showed that there was some evidence that the relationship between these two variables can be described by a monotonic, in this case, increasing function. To have a better understanding of their potential correalation, we also produced two plots. The first plot was produced using the original scale of the two variables. The second one was drawn after taking logrithm to base 10 of both of them. The two plots also demostrated some weak positive correlation between the two variables.<br />

```{r}
df_4 = df_commits %>%
  select(commit_sha, message, change_in_lines) %>%
  distinct() %>%
  select(message, change_in_lines) 
  
  cor.test(nchar(df_4$message), df_4$change_in_lines, method = "pearson")
  cor.test(nchar(df_4$message), df_4$change_in_lines, method = "spearman")
  ggplot(data = df_4) + geom_point(aes(x = nchar(message), y = change_in_lines), alpha = 0.25, color = "blue") + labs(x="Length of message",y="Changes in line")+ggtitle("Plot of Change_in_lines and message length")
  ggplot(data = df_4 ) + geom_point(aes(x = nchar(message), y = change_in_lines), alpha = 0.25, color = "blue") + 
    scale_y_log10() + scale_x_log10() +labs(x="log(Length of message)",y="log(Changes in line)")+ggtitle("Plot of log(Change_in_lines) and log(message length)")
```


5. Other than Romain Francois, which three contributors have worked on the most different files in the repository? In order words, who has the greatest number of unique file names among their commits.

Explanation to problem 5:<br />
For this problem, we selected `author_name` and `file_name` from `df_commits` at the beginning. Then we used `distinct()` to remove repeated files. Next, we counted the number of files each author worked on and excluded Romain Francois from the name list. Finally, we used `top_n()` to include tide values and derived the final results.<br />

```{r}
df_commits %>%
  select(author_name, file_name) %>%
  distinct() %>%
  group_by(author_name) %>%
  summarise(number_of_files_worked_on=n()) %>%
  arrange(-number_of_files_worked_on) %>%
  filter(author_name != "romainfrancois") %>%
  top_n(3, number_of_files_worked_on) %>%
  print(n = Inf)
```